\LARGE \textbf{Constrained Geometric Attitude Control on \( \SO \)}

### 2016-12-05 manuscript.tex
Add a subsection for the discussion about euler angles
Add time varying disturbance simulation

% Discuss the benefits of using this geometric representation as compared to Euler angles or quaternions
Attitude parameterizations, such as Euler angles and Quaternions, are frequently used in the aerospace and astrodynamics communities~\cite{vallado2001,bate1971,wertz1978}.
For example, Euler angle sequences are frequently used to describe the transformation between a variety of reference frames used to describe the position and orientation of the orbit of Earth satellites~\cite{vallado2001}.
In addition, quaternions were used during the operation of Skylab and the NASA Space Shuttle~\cite{hughes2004}.
However, the choice of attitude parameterization plays a critical role in control design and the resulting motion of the system.

% singularities and the problems involved with using Euler angles
Euler angle sequences are a minimum, three-parameter set of angles which describe the transformation between two reference frames.
Euler angles represent any general rotation as a sequence of three intermediate rotations.
By convention, there are \num{24} possible Euler angle sequences for any given rotation.
In addition, Euler angles are a minimum representation, as only three angles, and the associated sequence, are required to describe the three angular degrees of freedom of the rigid body.
However, there is great ambiguity in the representation of the attitude as there are many equivalent Euler angle sequences for a given attitude of the system.
Therefore, great care must be taken in the control system design to ensure that a consistent sequence is used. 
Furthermore, it has been shown that no minimal attitude representation can describe orientations both globally and without singularities~\cite{hughes2004,bhat2000}.
These singularities can cause significant difficulties during control design and hardware implementation.
\end{multicols}
\begin{figure} 
  \centering 
    \subfigure[Configuration error \( \Psi \)\label{fig:Psi_adapt}]{\includegraphics[width=0.4\columnwidth]{Psi_adapt} }~
    \subfigure[Angle to each constraint \label{fig:con_angles}]{\includegraphics[width=0.4\columnwidth]{constraint_angles} }\\
    \subfigure[Disturbance estimate \( \bar \Delta \) components \label{fig:delta_adapt} ]{\includegraphics[width=0.4\columnwidth]{Delta_adapt}}~
    \subfigure[Attitude trajectory \label{fig:cad_adapt}]{\includegraphics[trim={10cm 0 10cm 0},clip,width=0.4\columnwidth]{cad_adapt} }
  \caption{Attitude stabilization with adaptive update law}
  \label{fig:adapt} 
\end{figure}
\begin{multicols}{2}
To demonstrate the effect of the kinematic singularities inherent with Euler angles we will represent the attitude of the body fixed reference frame, \( \vecbf{b}_i \), with respect to the inertial frame, \( \vecbf{e}_i\), in terms of the \( 3-1-3\) Euler angle sequence.
More explicitly, this corresponds to the rotation sequence \( \theta_1 \vecbf{b}_3 , \theta_2 \vecbf{b}_1, \theta_3 \vecbf{b}_3 \).
The rotation matrix, \( R(\theta_1, \theta_2, \theta_3) \), corresponding to this sequence is 
\begin{align}\label{eq:euler313}
    \begin{bmatrix}
        -s_1 c_2 s_3 + c_3 c_1 & -s_1 c_2 c_3 - s_3 c_1 & s_1s_2 \\
        c_1 c_2 s_3 + c_3 s_1 & c_1 c_2 c_3 - s_3 s_1 & - c_1 s_2 \\
        s_2 s_3 & s_2 c_3 & c_2
    \end{bmatrix} ,
\end{align}
where \( s_i, c_i \) represent \( \sin \theta_i, \cos \theta_i \) for \( i = \braces{1,2,3}\).
Using this representation that kinematic differential equations for the associated Euler angles are given as
\begin{align}\label{eq:euler313_diff}
    \begin{bmatrix}
        \dot{\theta}_1 \\ \dot{\theta}_2 \\ \dot{\theta}_3 
    \end{bmatrix}
    =
    \begin{bmatrix}
        \slfrac{\parenth{\Omega_1 s_3 + \Omega_2 c_3}}{s_2} \\
        \Omega_1 c_3 - \Omega_2 s_3 \\
        -\slfrac{\parenth{\Omega_1 s_3 + \omega_2 c_3}c_2}{s_2} + \Omega_3
    \end{bmatrix} .
\end{align}
From~\cref{eq:euler313_diff}, it is immediately clear that a singularity exists when \( \sin \theta_2 = 0 \) or \( \theta_2 = 0, \pm \pi \). 
In the vicinity of the singularity, the angular velocities of the Euler angles will tend to approach \( \pm \infty \) and the angular velocities will experience instantaneous sign changes.
Furthermore, all Euler angle sequences will exhibit a similar singularity at either \( \theta_2 = 0, \pm \pi \) or \( \theta_2 = \pm \frac{\pi}{2}, \pm \frac{3\pi}{2} \) and switching the sequence does not alleviate the issue.
As a result, Euler angles are not appropriate for systems which experience large angular rotations, such as those demonstrated in~\cref{fig:adapt}, or control systems which rely on the angular velocities \( \theta_i \).

% discuss the unwinding phenomenon with quaternions
\section{Experiment on Hexrotor UAV}\label{sec:experiment}

>>>>

\subsection{Attitude Parameterizations}
% Discuss the benefits of using this geometric representation as compared to Euler angles or quaternions
Attitude parameterizations, such as Euler angles and Quaternions, are frequently used in the aerospace and astrodynamics communities~\cite{vallado2001,bate1971,wertz1978}.
For example, Euler angle sequences are frequently used to describe the transformation between a variety of reference frames used to describe the position and orientation of the orbit of Earth satellites~\cite{vallado2001}.
In addition, quaternions were used during the operation of Skylab and the NASA Space Shuttle~\cite{hughes2004}.
However, the choice of attitude parameterization plays a critical role in control design and the resulting motion of the system.

% singularities and the problems involved with using Euler angles
Euler angle sequences are a minimum, three-parameter set of angles which describe the transformation between two reference frames.
Euler angles represent any general rotation as a sequence of three intermediate rotations.
By convention, there are \num{24} possible Euler angle sequences for any given rotation.
In addition, Euler angles are a minimum representation, as only three angles, and the associated sequence, are required to describe the three angular degrees of freedom of the rigid body.
However, there is great ambiguity in the representation of the attitude as there are many equivalent Euler angle sequences for a given attitude of the system.
Therefore, great care must be taken in the control system design to ensure that a consistent sequence is used. 
Furthermore, it has been shown that no minimal attitude representation can describe orientations both globally and without singularities~\cite{hughes2004,bhat2000}.
These singularities can cause significant difficulties during control design and hardware implementation.
\end{multicols}
\begin{figure} 
  \centering 
    \subfigure[Configuration error \( \Psi \)\label{fig:Psi_adapt}]{\includegraphics[width=0.4\columnwidth]{Psi_adapt} }~
    \subfigure[Angle to each constraint \label{fig:con_angles}]{\includegraphics[width=0.4\columnwidth]{constraint_angles} }\\
    \subfigure[Disturbance estimate \( \bar \Delta \) components \label{fig:delta_adapt} ]{\includegraphics[width=0.4\columnwidth]{Delta_adapt}}~
    \subfigure[Attitude trajectory \label{fig:cad_adapt}]{\includegraphics[trim={10cm 0 10cm 0},clip,width=0.4\columnwidth]{cad_adapt} }
  \caption{Attitude stabilization with adaptive update law}
  \label{fig:adapt} 
\end{figure}
\begin{multicols}{2}
To demonstrate the effect of the kinematic singularities inherent with Euler angles we will represent the attitude of the body fixed reference frame, \( \vecbf{b}_i \), with respect to the inertial frame, \( \vecbf{e}_i\), in terms of the \( 3-1-3\) Euler angle sequence.
More explicitly, this corresponds to the rotation sequence \( \theta_1 \vecbf{b}_3 , \theta_2 \vecbf{b}_1, \theta_3 \vecbf{b}_3 \).
The rotation matrix, \( R(\theta_1, \theta_2, \theta_3) \), corresponding to this sequence is 
\begin{align}\label{eq:euler313}
    \begin{bmatrix}
        -s_1 c_2 s_3 + c_3 c_1 & -s_1 c_2 c_3 - s_3 c_1 & s_1s_2 \\
        c_1 c_2 s_3 + c_3 s_1 & c_1 c_2 c_3 - s_3 s_1 & - c_1 s_2 \\
        s_2 s_3 & s_2 c_3 & c_2
    \end{bmatrix} ,
\end{align}
where \( s_i, c_i \) represent \( \sin \theta_i, \cos \theta_i \) for \( i = \braces{1,2,3}\).
Using this representation that kinematic differential equations for the associated Euler angles are given as
\begin{align}\label{eq:euler313_diff}
    \begin{bmatrix}
        \dot{\theta}_1 \\ \dot{\theta}_2 \\ \dot{\theta}_3 
    \end{bmatrix}
    =
    \begin{bmatrix}
        \slfrac{\parenth{\Omega_1 s_3 + \Omega_2 c_3}}{s_2} \\
        \Omega_1 c_3 - \Omega_2 s_3 \\
        -\slfrac{\parenth{\Omega_1 s_3 + \omega_2 c_3}c_2}{s_2} + \Omega_3
    \end{bmatrix} .
\end{align}
From~\cref{eq:euler313_diff}, it is immediately clear that a singularity exists when \( \sin \theta_2 = 0 \) or \( \theta_2 = 0, \pm \pi \). 
In the vicinity of the singularity, the angular velocities of the Euler angles will tend to approach \( \pm \infty \) and the angular velocities will experience instantaneous sign changes.
Furthermore, all Euler angle sequences will exhibit a similar singularity at either \( \theta_2 = 0, \pm \pi \) or \( \theta_2 = \pm \frac{\pi}{2}, \pm \frac{3\pi}{2} \) and switching the sequence does not alleviate the issue.
As a result, Euler angles are not appropriate for systems which experience large angular rotations, such as those demonstrated in~\cref{fig:adapt}, or control systems which rely on the angular velocities \( \theta_i \).
% discuss the unwinding phenomenon with quaternions

\subsection{Time-varying Disturbance}\label{ssec:time_varying}

The form of the uncertainty, given in~\cref{eqn:Wdot}, is commonly used in the adaptive control literature~\cite{lee2013b,ioannou2012}. 
A wide variety of realistic disturbances, such as gravitational gradients or malfunctioning thrusters, are accurately represented via this model. 
In addition, it is possible to represent the uncertainty of a time-varying inertia matrix as an equivalent external disturbance. 
For example, Euler's law gives the relationship for the rate of change of angular momentum
\begin{align*}
    M_{ext} = \dot{\vecbf{H}} = \dot{J} \vecbf{\Omega} + J \dot{\vecbf{\Omega}} .
\end{align*}
Using this, we can see that an instantaneous change in \( J \) is proportional to an external moment.
Finally, it has been shown that this adaptive control formulation is able to handle time-varying disturbances under some mild assumptions~\cite{ioannou2012}.

We demonstrate the ability to handle an uncertain time-varying disturbance via numerical example.
The system is identical to the one presented in~\Cref{sec:numerical_simulation}, however we modify the external disturbance. 
We model the external disturbance as the superposition of constant and time-varying terms as
\begin{align*}
    \Delta = \begin{bmatrix} 0.2 \\ 0.2 \\0.2 \end{bmatrix} + 0.02 \begin{bmatrix} \sin 9 t \\ \cos 9 t \\ \frac{1}{2} \parenth{\sin 9t + \cos 9t}\end{bmatrix} \si{\newton\meter}.
\end{align*}
We define a constraint in the inertial frame as \( v = [\frac{1}{\sqrt{2}}, \frac{1}{\sqrt{2}}, 0]^T \) with \( \theta = \ang{12} \).
The initial state is defined as \(R(0) = \exp( \frac{\pi}{2} \hat{e}_3) \), while the desired state is \(R_d =I \).
The goal is to rotate the vehicle about the \( e_3 \) axis while avoiding the obstacle and compensating for the time-varying disturbance.

\Cref{fig:tv} demonstrates the ability for the adaptive controller, presented in~\cref{prop:adaptive_control}, to handle time-varying disturbances.
We can see in~\cref{fig:Delta_tv} that that estimate \(\bar \Delta \) accurately tracks the true disturbance after approximately~\SI{5}{\second}.
In addition,~\cref{fig:con_angle_tv} shows that the constraint is never verified as the angle between the body-fixed sensor \( r \) and the constraint \( v \) is greater than \SI{12}{\degree} over the entire attitude maneuver.
\end{multicols}
\begin{figure}[H]
    \centering 
    \subfigure[{Configuration error \( \Psi \)} \label{fig:Psi_tv} ]{\includegraphics[width=0.3\columnwidth]{Psi_tv} }~
    \subfigure[{Angle to constraint } \label{fig:con_angle_tv}]{\includegraphics[width=0.3\columnwidth]{figures/constraint_angle_tv.pdf} }~
    \subfigure[{Disturbance estimate \(\bar\Delta\) components} \label{fig:Delta_tv}]{\includegraphics[width=0.3\columnwidth]{figures/delta_tv} }
    \caption{Time-varying external disturbance simulation}
    \label{fig:tv} 
\end{figure}
\begin{multicols}{2}
\section{Experiment on Hexrotor UAV}\label{sec:experiment}
